Novel lattice Boltzmann method for simulation of strongly shear thinning viscoelastic fluids

### 2024-06-18 assets/responseToReview.txt

-Is the relaxation time τ_r=1 adopted in numerical simulations?

Yes, see 4.1

-Why is the finite-volume method adopted for polymer stress? Please give some details on the finite-volume method. In addition, it is known that in the lattice Boltzmann method, the square or cubic lattice is adopted, and thus the finite-difference method seems a better choice?

Makes boundary conditions trivial (flux = 0); CTU (which is a FV scheme) has been reported to increase stability [Kuron]
Added some more details to 4.2 how we use FV here.
Finite difference is a lot harder to implement (boundary conditions), has a lot more loose ends to tie (to assure conservation for example) and is less stable.

@@

(1)The introduction is too simple. The development of LBM for simulating viscoelastic fluids is not reviewed thoroughly.

# TODO this is physics and not a history lecture -> Delegated to Stephan

(2) The numerical method should be given in detail and in a logical way. Section 4.1 does not given in detail of the LBM. Section 4.2 tries to show the finite-volume method for polymer stress but only presents the LBM for the flow field, which is incorrect. The finite volume method for the polymer stress is actually missing.

Added additional LBM section
You seem to suggest, that the stated LBM equation is incorrect. Please elaborate.
Added some more details to 4.2 how we use FV here.

### 2024-06-18 NovelLBM.tex
FV wording

The Lattice-Boltzmann method (LBM) has become a standard tool in numerical fluid mechanics \cite{Kruger_book}, and we therefore \RR{only mention the key choices here. Further details can be found in the appendix~\ref{app:LBM} or in literature.}
We use a D3Q19 lattice in full three dimensions together with a single relaxation time collision operator.
Our implementation is based on FluidX3D which has been extensively used and validated in earlier publications \cite{Lehmann.20229da}.
The core of the algorithm is implemented in OpenCL and thus capable of running on GPUs from different vendors.
Non-periodic boundaries are modeled using the half-way bounce back algorithm. To model fixed inflow/outflow, bounce back boundaries using a Dirichlet boundary condition as described by Krüger \textit{et al.} \cite{Kruger_book} are used.

@@

The spatio-temporal dynamics of the polymer stress $\ma{\tau}$ is given by equation~(\ref{eq:polymerStress}) in the form of an advection-reaction equation.
We solve this equation by a finite-volume scheme running in parallel (with identical time step) to the LBM solver.
Advection is handled using the corner transport upwind scheme \cite{kuron2021}.
At the walls, a zero-flux boundary condition for the stress is used.
The source terms are integrated using the Euler method.

>>>>

The Lattice-Boltzmann method (LBM) has become a standard tool in numerical fluid mechanics \cite{Kruger_book}, and we therefore \RR{only mention the key choices here. Further details can be found in the appendix~\ref{app:LBM} or in literature.}
We use a D3Q19 lattice in full three dimensions together with a single relaxation time collision operator.
Our implementation is based on FluidX3D which has been extensively used and validated in earlier publications \cite{lehmann2021, Lehmann.20229da}.
The core of the algorithm is implemented in OpenCL and thus capable of running on GPUs from different vendors.
Non-periodic boundaries are modeled using the half-way bounce back algorithm. To model fixed inflow/outflow, bounce back boundaries using a Dirichlet boundary condition as described by Krüger \textit{et al.} \cite{Kruger_book} are used.

@@

The spatio-temporal dynamics of the polymer stress $\ma{\tau}$ is given by equation~(\ref{eq:polymerStress}) in the form of an advection-reaction equation.
We solve this equation by a finite-volume scheme running in parallel (with identical time step) to the LBM solver.\RR{ For this, the simulation volume is subdivided in $1\times1\times1$ (in LU) cubes centered around each lattice node. The average value of $\ma{\tau}$ within these volumes is takes as the value at the lattice node.}
Advection is handled using the corner transport upwind scheme \cite{kuron2021}.\RR{ The remaining derivatives are calculated using finite differences.}
At the walls\RR{, which coincide with the surfaces of the finite volumes}, a zero-flux boundary condition for the stress is used.
The source terms are integrated using the Euler method.
